---
title: "License Movement"
author: "cslovas"
format: 
  html:
    toc: true
    self-contained: true
editor: visual
---

```{r}
#| label: read in geocoded garfo data
#| echo: false
#| message: false

library(tidyverse)
library(gmRi)
library(here)
library(matrixStats)

garfo_geo    <- read.csv(file = here("Outputs", "GARFO_geocoded.csv"))
ny_nj        <- read.csv(file = here("Outputs", "geo_ny_nj.csv"))
species_list <- c("Atlantic menhaden", "Black sea bass", "Bluefish", "Clearnose skate", "Horseshoe crab", "Longfin squid", "Rosette skate", "Scup", "Striped bass", "Summer flounder", "Tilefish", "Weakfish")

```

## Tracking latitudinal movement of GARFO licenses

By rounding latitude to 1/10 of a degree, we can calculate the center of gravity/occurrence of a license type. To do this, we weight the average latitude by the total number of a particular license type associated with that latitude over time.

```{r}
#| label: grouped by target species 
#| echo: false
#| message: false

garfo_geo <- garfo_geo %>% full_join(ny_nj)


species_list <- as_tibble(species_list) %>% rename("species" = "value") 

license_list <- garfo_geo %>%
  ungroup() %>%
  pivot_longer(cols = BLACK_SEA_BASS_1:TILEFISH_D, names_to = "license", values_to = "count") %>%
  select(license) %>% 
  distinct() %>%
  mutate(category = NA)

license_list$category[str_starts(license_list$license, "BLACK_SEA_BASS")] = "Black sea bass"
license_list$category[str_starts(license_list$license, "BLUEFISH")] = "Bluefish"
license_list$category[str_starts(license_list$license, "DOGFISH")] = "Dogfish"
license_list$category[str_starts(license_list$license, "GEN_CAT")] = "Scallop"
license_list$category[str_starts(license_list$license, "HERRING")] = "Herring"
license_list$category[str_starts(license_list$license, "HMS_")] = "Squid"
license_list$category[str_starts(license_list$license, "LOBSTER")] = "Lobster"
license_list$category[str_starts(license_list$license, "MONKFISH")] = "Monkfish"
license_list$category[str_starts(license_list$license, "MULTISPECIES_")] = "Multispecies"
license_list$category[str_starts(license_list$license, "QUAHOG")] = "Quahog"
license_list$category[str_starts(license_list$license, "RED")] = "Red crab"
license_list$category[str_starts(license_list$license, "SCUP")] = "Scup"
license_list$category[str_starts(license_list$license, "SEA_SCALLOP")] = "Sea scallop"
license_list$category[str_starts(license_list$license, "SKATE")] = "Skate"
license_list$category[str_starts(license_list$license, "SQUID_")] = "Squid/Mackerel/Butterfish"
license_list$category[str_starts(license_list$license, "SUMMER")] = "Summer flounder"
license_list$category[str_starts(license_list$license, "SURF")] = "Surf clam"
license_list$category[str_starts(license_list$license, "TILE")] = "Tilefish"

write_csv(license_list, here("Data", "garfo_license_list.csv"))
```

### Black sea bass

Black sea bass is used as an example, as there are only two license types and a distinguishable trend of northward (poleward) latitudinal movement.

```{r}
#| label: restructuring dataframe
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 7 

garfo_geo %>%
  ungroup() %>%
  pivot_longer(cols = BLACK_SEA_BASS_1:TILEFISH_D, names_to = "license", values_to = "count") %>%
  filter(!count == 0) %>%
  left_join(license_list) %>%
  group_by(VP_NUM, category) %>%
  mutate(lat = round(lat, digits = 1),
         long = round(long, digits = 1)) %>%
  filter(lat >= 30) %>%
  group_by(AP_YEAR, category, lat) %>%
  summarise(count = sum(count)) %>%
  group_by(AP_YEAR, category) %>%
  summarise(center_lat = weightedMean(lat, w = count)) %>%
  distinct() -> center_lats

center_lats %>% 
  filter(category == "Black sea bass") %>%
  # filter(str_starts(license, "BLACK_SEA_BASS")) %>% 
  ggplot() +
  geom_point(aes(x = AP_YEAR, y = center_lat)) +  #, color = license)) +
  xlab("Year") + ylab("Center of Latitude") +
  ylim(c(37,42)) +
  scale_color_gmri() + 
  theme_gmri()
 
```

### Bluefish

```{r}
#| label: bluefish plot
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 7 


center_lats %>% 
  filter(category == "Bluefish") %>%
  #filter(str_starts(license, "BLUEFISH")) %>% 
  ggplot() +
  # geom_line(aes(x = AP_YEAR, y = center_lat)) +
  geom_point(aes(x = AP_YEAR, y = center_lat)) +# color = license)) +
  xlab("Year") + ylab("Center of Latitude") + 
  # facet_wrap(~license, ncol = 3) + 
  # ylim(c(35,45)) +
  # scale_color_gmri() + 
  theme_gmri(legend.position = "none",
             strip.background = element_rect(fill = "lightgray"),
             strip.text = element_text(color = "black"))
```

### Multispecies

```{r}
#| label: multispecies
#| echo: false
#| message: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 7 

center_lats %>% 
  filter(category == "Multispecies") %>%
  ggplot() +
  # geom_line(aes(x = AP_YEAR, y = center_lat)) +
  geom_point(aes(x = AP_YEAR, y = center_lat)) + # color = license)) +
  xlab("Year") + ylab("Center of Latitude") + 
  theme_gmri(legend.position = "none",
             strip.background = element_rect(fill = "lightgray"),
             strip.text = element_text(color = "black"))

```

### Squid, mackerel, and butterfish

```{r}
#| label: squid
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 7 

center_lats %>% 
  filter(category == "Squid/Mackerel/Butterfish") %>%
  # filter(str_starts(license, "SQUID")) %>% 
  ggplot() +
  geom_point(aes(x = AP_YEAR, y = center_lat))+ #, color = license)) +
  xlab("Year") + ylab("Center of Latitude")  +
  # facet_wrap(~license, ncol = 3) + 
  ylim(c(35,45)) +
  scale_color_gmri() + 
  theme_gmri(legend.position = "none",
             strip.background = element_rect(fill = "lightgray"),
             strip.text = element_text(color = "black"))
```

### Skate

```{r}
#| label: skate
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 7 


center_lats %>% 
  filter(category == "Skate") %>%
  ggplot() +
  geom_point(aes(x = AP_YEAR, y = center_lat)) +
  xlab("Year") + ylab("Center of Latitude") + 
  # facet_wrap(~license, ncol = 3) + 
  ylim(c(35,45)) +
  scale_color_gmri() + 
  theme_gmri(legend.position = "none",
             strip.background = element_rect(fill = "lightgray"),
             strip.text = element_text(color = "black"))
```

### Scup

```{r}
#| label: scup
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 7 


center_lats %>% 
  filter(category == "Scup") %>%
  ggplot() +
  geom_point(aes(x = AP_YEAR, y = center_lat)) +
  xlab("Year") + ylab("Center of Latitude") + 
  # facet_wrap(~license, ncol = 3) + 
  ylim(c(35,45)) +
  scale_color_gmri() + 
  theme_gmri(legend.position = "none",
             strip.background = element_rect(fill = "lightgray"),
             strip.text = element_text(color = "black"))
```

## Geographic spread

```{r}
#| label: maps
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 7
#| fig-width: 10 

library(sf)
library(rnaturalearth)

world <- ne_countries(returnclass = "sf")

garfo_geo %>%
  ungroup() %>%
  pivot_longer(cols = BLACK_SEA_BASS_1:TILEFISH_D, names_to = "license", values_to = "count") %>%
  filter(!count == 0) %>%
  left_join(license_list) %>%
  filter(category == "Multispecies") %>% 
  mutate(decade = 10*AP_YEAR %/% 10) -> multispecies_map

ggplot() +
  geom_sf(data = world) +
  geom_point(data = multispecies_map, aes(x = long, y = lat, color = decade)) +
  coord_sf(xlim = c(-78,-66), ylim = c(34, 47)) +
  scale_x_continuous(breaks = c(-76, -72, -68)) +
  scale_y_continuous(breaks = c(34, 38, 42, 46)) +
  ggtitle("Multispecies licenses") + xlab("Lat") + ylab("Lon") +
  facet_wrap(~decade, nrow = 1) +
  theme_gmri(strip.background = element_rect(fill = "transparent", color = "black", linewidth = 0.7),
             strip.text = element_text(color = "black"),
             # legend.title = element_text(size = 15, face = "bold"),
             # legend.text = element_text(size = 15),
             legend.position = "none",
             plot.title = element_text(size = 15, face = "bold"),
             plot.subtitle = element_text(size = 13),
             axis.title = element_text(size = 12, face = "bold"),
             axis.text = element_text(size = 12),
             panel.border = element_rect(color = "black", linetype = 1, linewidth = 0.5))
```

There's no clear trend at this scale of license distribution changing. In the center of gravity analysis, there is a strong southward trend of multispecies groundfish licenses. We will run a GINI analysis later but first will plot the raw latitude data as boxplot to understand the spread of these license categories. It might be useful to compare the groundfish landings trends to the license trend so see if there are correlative shifts.

```{r}
#| label: boxplot
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 7
#| fig-width: 10 
ggplot(data = multispecies_map) +
  geom_boxplot(aes(x = AP_YEAR, y = lat, group = AP_YEAR), outlier.alpha = 0.2) +
  ggtitle("Multispecies licenses", "Boxplot of license spread") +
  ylab("Latitude") + xlab("Year") +
  ylim(c(38,46)) +
  theme_gmri()

```

There is a very subtle trend of the median moving southwards, likely being pulled towards all the southern outliers, which seem to outnumber northward outliers.

```{r}
#| label: boxplots for all
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 7
#| fig-width: 10 

garfo_geo %>%
  ungroup() %>%
  pivot_longer(cols = BLACK_SEA_BASS_1:TILEFISH_D, names_to = "license", values_to = "count") %>%
  filter(!count == 0) %>%
  left_join(license_list) %>%
  mutate(decade = 10*AP_YEAR %/% 10) %>% 
  group_by(category) %>%
  filter(!category == "Multispecies") %>% # already above
  nest() %>%
  mutate(boxplot = map2(data, category, function(x,y){
    ggplot(data = x) +
      geom_boxplot(aes(x = AP_YEAR, y = lat, group = AP_YEAR), outlier.alpha = 0.2) +
      ggtitle(y, "Boxplot of license spread") +
      ylab("Latitude") + xlab("Year") +
      ylim(c(36,46)) +
      theme_gmri()
  })) -> boxplots

gridExtra::marrangeGrob(boxplots$boxplot, layout_matrix = matrix(1:1, ncol = 1, nrow = 1, byrow = TRUE), top = NULL)
```

```{r}
#| label: maps for all
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 7
#| fig-width: 10 

garfo_geo %>%
  ungroup() %>%
  pivot_longer(cols = BLACK_SEA_BASS_1:TILEFISH_D, names_to = "license", values_to = "count") %>%
  filter(!count == 0) %>%
  left_join(license_list) %>%
  mutate(decade = 10*AP_YEAR %/% 10) %>% 
  group_by(category) %>%
  filter(!category == "Multispecies") %>% # already above
  nest() %>%
  mutate(maps = map2(data, category, function(x,y){
    ggplot() +
      geom_sf(data = world) +
      geom_point(data = x, aes(x = long, y = lat, color = decade)) +
      coord_sf(xlim = c(-78,-66), ylim = c(34, 47)) +
      scale_x_continuous(breaks = c(-76, -72, -68)) +
      scale_y_continuous(breaks = c(34, 38, 42, 46)) +
      ggtitle(y) + xlab("Lat") + ylab("Lon") +
      facet_wrap(~decade, nrow = 1) +
      theme_gmri(strip.background = element_rect(fill = "transparent", color = "black", linewidth = 0.7),
             strip.text = element_text(color = "black"),
             legend.position = "none",
             plot.title = element_text(size = 15, face = "bold"),
             plot.subtitle = element_text(size = 13),
             axis.title = element_text(size = 12, face = "bold"),
             axis.text = element_text(size = 12),
             panel.border = element_rect(color = "black", linetype = 1, linewidth = 0.5))
  })) -> maps

gridExtra::marrangeGrob(maps$maps, layout_matrix = matrix(1:1, ncol = 1, nrow = 1, byrow = TRUE), top = NULL)
```

New Jersey appears to be missing from all of these...
