---
title: "Squid"
author: "cslovas"
format: 
  html:
    toc: true
    self-contained: true
editor: visual
---

## What is happening with squid

#### Helpful information on longfin and shortfin squid

-   [Atlantic States Marine Fisheries Commission](https://www.asmfc.org/species/black-sea-bass)
-   [NOAA Fisheries](https://www.fisheries.noaa.gov/species/black-sea-bass)
-   [Federal permits](https://www.fisheries.noaa.gov/permit/mackerel-squid-and-butterfish)

```{r}
#| label: load data and dependencies 
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
# devtools::install_github("https://github.com/gulfofmaine/gmRi")
library(gmRi)
library(here)
library(grid)
library(gomfish)
library(matrixStats)
library(rnaturalearth)

garfo_clean    <- read.csv(here("Data", "all_GARFO_data.csv"))
garfo_landings <- read.csv(here("Data", "FOSS_landings_squid.csv"), header = TRUE)
garfo_licenses <- read.csv(here("Data", "garfo_license_list.csv"))
maine_licenses <- ind_lic_portfolio
maine_codes    <- read.csv(here("Data", "updated_license_codes.csv"))

```

```{r}
#| label: load trawl data
#| echo: false
#| message: false
#| warning: false

# Load NEFSC Bottom Trawl Survey data ####
clean_survey <- gmri_survdat_prep(
  survdat_source ="most recent",
  box_location ="cloudstorage"
)

clean_survey  <- clean_survey %>% 
  distinct(est_year, survey_area, stratum, tow, est_towdate, season, comname, catchsex, .keep_all = T) %>%
  filter(!season %in% c("Summer", "Winter") & est_year >= 1970) %>%
  group_by(est_year, survey_area, stratum, tow, est_towdate, season, 
           avgdepth, surftemp, bottemp, decdeg_beglat, decdeg_beglon, comname, abundance) %>% 
  summarise(biomass_kg = sum(biomass_kg, na.rm = T), .groups = "drop") 

# Weight by biomass
grouped_center_bio <- function(clean_survey, ...){
  clean_survey %>% 
    group_by(comname, ...) %>% 
    summarise(
      # Un-weighted averages
      total_biomass   = sum(biomass_kg),
      avg_biomass     = mean(biomass_kg),
      biomass_sd      = sd(biomass_kg),
      # All below are weighted by biomass
      avg_depth       = weightedMean(avgdepth, w = biomass_kg, na.rm = T),
      avg_bot_temp    = weightedMean(bottemp, w = biomass_kg, na.rm = T),
      avg_sur_temp    = weightedMean(surftemp, w = biomass_kg, na.rm = T),
      avg_lat         = weightedMean(decdeg_beglat, w = biomass_kg, na.rm = T),
      avg_lon         = weightedMean(decdeg_beglon, w = biomass_kg, na.rm = T),
      depth_sd        = weightedSd(avgdepth, w = biomass_kg, na.rm = T),
      temp_sd         = weightedSd(bottemp, w = biomass_kg, na.rm = T),
      lat_sd          = weightedSd(decdeg_beglat, w = biomass_kg, na.rm = T),
      lon_sd          = weightedSd(decdeg_beglon, w = biomass_kg, na.rm = T),
      .groups = "drop") 
}

weighted_data <- grouped_center_bio(clean_survey, est_year, season) %>%
  mutate(decade = 10*est_year %/% 10)
```

```{r}
#| label: squid
#| echo: false
#| warning: false
#| message: false
#| fig-align: center

biomass <- weighted_data %>% 
  filter(comname %in% c("northern shortfin squid", "longfin squid") & est_year >= 1970 & !season %in% c("Winter", "Summer")) 

landings <- garfo_landings %>%
  filter(Year >= 1970) %>%
  select(Year, State, NMFS.Name, Pounds, Dollars, Confidentiality) %>% 
  mutate(Pounds  = parse_number(Pounds),
         Dollars = parse_number(Dollars))

fed_license <- garfo_clean %>% 
  select(AP_YEAR, AP_NUM, VP_NUM, PPORT, PPST, starts_with("SQUID_")) %>% # leaving out HMS_Squid
  pivot_longer(cols = 6:17, names_to = "license", values_to = "count") %>% 
  filter(PPST %in% c("ME","NH","MA","RI","CT","NY","NJ","DE","MD","VA","NC")) %>%  
  filter(!count == 0)

```

## Distributions

### Average biomass and center of biomass based on NEFSC bottom trawl survey

```{r}
#| label: biomass
#| echo: false
#| message: false
#| warnging: false
#| fig-align: center

ggplot(data = biomass) +
  geom_line(aes(x = est_year, y = avg_lat, color = season)) +
  theme_gmri()+
  scale_color_gmri() +
  facet_wrap(~comname) +
  ggtitle("Seasonal center of biomass") +
  xlab("Year") + ylab("Biomass-weighted average latitude")

biomass %>%
  group_by(comname, est_year) %>% 
  summarise(avg_lat = mean(avg_lat)) %>% 
  ggplot() +
  geom_line(aes(x = est_year, y = avg_lat)) + 
  geom_smooth(aes(x = est_year, y = avg_lat), method = "lm", se = FALSE, color = "#07a3b7", linetype = 2, alpha = 0.7) +
  ylab("Biomass-weighted average latitude") +
  facet_wrap(~comname) +
  xlab("Year") +
  ggtitle("Annual center of biomass") +
  theme_gmri()

biomass %>% 
  group_by(est_year, season, comname) %>%
  summarise(total_biomass = sum(total_biomass)) %>% 
  ggplot()+
  geom_col(aes(x = est_year, y = total_biomass, fill = season), position = "dodge") + #fill = "#07a3b7") +
  scale_fill_gmri(palette = "main") +
  theme_gmri() +
  facet_wrap(~comname) +
  ggtitle("Annual surveyed biomass") +
  ylab("Biomass (kg)") +
  xlab("Year")

biomass %>% 
  group_by(est_year, comname) %>%
  summarise(annual_biomass = mean(total_biomass)) %>% 
  ggplot()+
  geom_line(aes(x = est_year, y = annual_biomass)) + #fill = "#07a3b7") +
  scale_fill_gmri(palette = "main") +
  theme_gmri() +
  facet_wrap(~comname) +
  ggtitle("Annual mean surveyed biomass") +
  ylab("Biomass (kg)") +
  xlab("Year")


```
Nothing with shortfin is really jumping out to me, so I'll move forward with just longfin squid (unless Emily tells me otherwise). 
### Leading and trailing edges

```{r}
#| label: edges
#| echo: false
#| message: false
#| warning: false
#| fig-align: center

clean_survey %>% 
  filter(comname == "longfin squid" & est_year >= 1970) %>%
  group_by(est_year) %>%  
  summarise(
    `10%`  = Hmisc::wtd.quantile(decdeg_beglat, weights = biomass_kg, probs = 0.10, na.rm = T),
    `90%`  = Hmisc::wtd.quantile(decdeg_beglat, weights = biomass_kg, probs = 0.90, na.rm = T)) %>% 
  pivot_longer(cols = 2:3, names_to = "quantile", values_to = "lat") %>% 
  left_join(
    clean_survey %>% 
      filter(est_year >= 1970) %>%
      group_by(est_year) %>%  
      summarise(
        `10%`  = Hmisc::wtd.quantile(decdeg_beglon, weights = biomass_kg, probs = 0.10, na.rm = T),
        `90%`  = Hmisc::wtd.quantile(decdeg_beglon, weights = biomass_kg, probs = 0.90, na.rm = T)) %>%
     pivot_longer(cols = 2:3, names_to = "quantile", values_to = "lon")
  ) -> quantiles

quantiles %>% 
  group_by(quantile) %>% 
   mutate(rmean_lat = zoo::rollapplyr(lat, width = 5, FUN = mean, align = "center", partial = T),
          rmean_lon = zoo::rollapplyr(lon, width = 5, FUN = mean, align = "center", partial = T)) %>%
  mutate(across(where(is.numeric), round, 4)) -> quantiles

ggplot(quantiles) +
  geom_line( aes(x = est_year, y = rmean_lat, color = quantile)) +
  geom_smooth(aes(x = est_year, y = rmean_lat, color = quantile), method = "lm", linetype = 2, linewidth = 0.5) +
  scale_color_gmri()+
  theme_gmri(plot.subtitle = element_text(size = 10)) +
  ggtitle("Longfin squid leading and trailing edge", subtitle = "5-year rolling mean latitude") +
  xlab("Year") + ylab("Latitude") +
  ylim(c(36,42))+
  guides(col = guide_legend(title = "Percentile"))

```

## Landings

### Federally reported landings

Dealer data provided by Greater Atlantic Regional Fisheries Office

```{r}
#| label: landings 
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-width: 11
#| fig-height: 9

landings <- landings %>% filter(NMFS.Name == "SQUID, LONGFIN LOLIGO")

landings$State_Abbr <- NA
landings$State_Abbr[landings$State == "MAINE"] = "ME"
landings$State_Abbr[landings$State == "NEW HAMPSHIRE"] = "NH"
landings$State_Abbr[landings$State == "MASSACHUSETTS"] = "MA"
landings$State_Abbr[landings$State == "RHODE ISLAND"] = "RI"
landings$State_Abbr[landings$State == "CONNECTICUT"] = "CT"
landings$State_Abbr[landings$State == "NEW YORK"] = "NY"
landings$State_Abbr[landings$State == "NEW JERSEY"] = "NJ"
landings$State_Abbr[landings$State == "DELAWARE"] = "DE"
landings$State_Abbr[landings$State == "MARYLAND"] = "MD"
landings$State_Abbr[landings$State == "VIRGINIA"] = "VA"
landings$State_Abbr[landings$State == "NORTH CAROLINA"] = "NC"

landings$Coast <- factor(landings$State_Abbr, levels = c("ME", "NH", "MA", "RI", "CT", "NY", "NJ", "DE", "MD", "VA", "NC" ))

landings %>%
  group_by(State, Year, Coast) %>% 
  summarise(total_volume = sum(Pounds)/1000000) %>%
  ggplot() +
  geom_col(aes(x = Year, y = total_volume), fill = "#07a3b7", alpha = 0.7) +
  facet_wrap(~Coast, ncol = 3, scales = "free_y") +
  ylim(c(0,NA)) +
  ggtitle("Landed volume") +
  ylab("Total landings (million lbs)") +
  xlab("Year") +
  theme_gmri(strip.background = element_rect(fill = "transparent"),
              strip.text = element_text(color = "black")) 

landings %>%
  group_by(State, Year, Coast) %>% 
  summarise(total_value = sum(Dollars)/1000000) %>%
  ggplot() +
  geom_line(aes(x = Year, y = total_value), color = "#057872") +
  facet_wrap(~Coast, ncol = 3, scales = "free_y") +
  ylim(c(0, NA)) +
  ggtitle("Landed value") +
  ylab("Total landings (million $USD)") +
  xlab("Year") +
  theme_gmri(strip.background = element_rect(fill = "transparent"),
              strip.text = element_text(color = "black")) 

# Total by state
landings %>% 
  drop_na() %>% 
  group_by(Year, State, Coast) %>% 
  summarise(total_volume = sum(Pounds)/1000000) %>% 
ggplot() +
  geom_col(aes(x = Year, y = total_volume, group = Coast, fill = Coast), color = "white", linewidth = 0.25) +
  ggtitle("Total landings by state") +
  xlab("Year") + ylab("Landed pounds (million)") +
  guides(fill = guide_legend("States")) +
  scale_fill_gmri() +
  theme_gmri() 
```

### Ex-vessel price of black sea bass

Despite a general upward trend in landed volume and value, the price per pound has dropped significantly in all regions. Maine is leftout due to lack of data.

```{r}
#| label: ex-vessel price
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 6
#| fig-width: 9

ex_vessel <- landings %>% 
  mutate(ex.vessel = Pounds/Dollars) 

ex_vessel %>% 
  group_by(State, Year, Coast) %>% 
  summarise(avg.ex.vessel = mean(ex.vessel)) %>% 
  # filter(!State == "MAINE") %>% # major outlier
  ggplot() + 
  geom_line(aes(x = Year, y = avg.ex.vessel, color = Coast)) +
  ggtitle("Average ex-vessel price") + 
  xlab("Year") + ylab("Price per pound (USD)") + 
  guides(color = guide_legend("States")) +
  theme_gmri() +
  scale_color_gmri()  
```
## Licensing

### Federal Licenses

Squid are managed under on fishery management plan (FMP) with mackerel and butterfish and are co-licensed with the same species. 

-   **Category 1A-C**: Limited access longfin squid (moratorium)

-   **Category 2**: Squid/Mackerel/Butterfish Charter Party

-   **Category 3**: Squid/Butterfish Incidental Catch

-   **Category 4**: Atlantic Mackerel

-   **Category 5**: Limited access Illex squid (moratorium)

-   **Category 6**: Limited access butterfish (moratorium)

-   **Category T1-3**: Limited access mackerel (tier 1-3)

Again, given the focus on longfin squid, we will focus only on *Category 1A-C*, *Category 3*, and *Category 5* licenses. 

```{r}
#| label: federal licenses
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-width: 11
#| fig-height: 9

# arrange north to south
fed_license <- fed_license %>% 
  filter(license %in% c("SQUID_MACK_BUTTER_1A","SQUID_MACK_BUTTER_1B", "SQUID_MACK_BUTTER_1C", "SQUID_MACK_BUTTER_3", "SQUID_MACK_BUTTER_5"))

fed_license$category <- NA
fed_license$category[fed_license$license == "SQUID_MACK_BUTTER_1A"] = "1A"
fed_license$category[fed_license$license == "SQUID_MACK_BUTTER_1B"] = "1B"
fed_license$category[fed_license$license == "SQUID_MACK_BUTTER_1C"] = "1C"
fed_license$category[fed_license$license == "SQUID_MACK_BUTTER_3"]  = "3"
fed_license$category[fed_license$license == "SQUID_MACK_BUTTER_5"]  = "5"

fed_license$states <- factor(fed_license$PPST, levels = c("ME", "NH", "MA", "RI", "CT", "NY", "NJ", "DE", "MD", "VA", "NC" ))

fed_license %>% 
  group_by(PPST, AP_YEAR, license, states, category) %>% 
  summarise(total_count = sum(count)) %>%
  ggplot() +
  geom_col(aes(x = AP_YEAR, y = total_count, fill = category)) +
  facet_wrap(~states, scales = "free_y", ncol = 3) +
  ggtitle("Commercial harvest licenses") +
  xlab("Year") + ylab("Total issued") +
  scale_fill_gmri() +
  theme_gmri(strip.background = element_rect(fill = "transparent"),
              strip.text = element_text(color = "black")) 

```
It seems the majority of license types targeting squid are incidental catch. 