---
title: "Black sea bass"
author: "cslovas"
format: 
  html:
    toc: true
    self-contained: true
editor: visual
---

## What is happening with black sea bass?

Pulling together bottom trawl survey data, federal landings, and federal license dat, here we explore black sea bass access and exploitation.

```{r}
#| label: load data and dependencies 
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(gmRi)
library(here)
library(grid)
library(gomfish)
library(matrixStats)

garfo_clean    <- read.csv(here("Data", "all_GARFO_data.csv"))
garfo_landings <- read.csv(here("Data", "landings.csv"))
garfo_licenses <- read.csv(here("Data", "garfo_license_list.csv"))
maine_licenses <- ind_lic_portfolio
maine_codes    <- read.csv(here("Data", "updated_license_codes.csv"))

```

```{r}
#| label: load trawl data
#| echo: false
#| message: false
#| warning: false

# Load NEFSC Bottom Trawl Survey data ####
clean_survey <- gmri_survdat_prep(
  survdat_source ="most recent",
  box_location ="cloudstorage"
)

clean_survey  <- clean_survey %>% 
  distinct(est_year, survey_area, stratum, tow, est_towdate, season, comname, catchsex, .keep_all = T) %>%
  group_by(est_year, survey_area, stratum, tow, est_towdate, season, 
           avgdepth, surftemp, bottemp, decdeg_beglat, decdeg_beglon, comname, abundance) %>% 
  summarise(biomass_kg = sum(biomass_kg, na.rm = T), .groups = "drop")

# Weight by biomass
grouped_center_bio <- function(clean_survey, ...){
  clean_survey %>% 
    group_by(comname, ...) %>% 
    summarise(
      # Un-weighted averages
      total_biomass   = sum(biomass_kg),
      avg_biomass     = mean(biomass_kg),
      biomass_sd      = sd(biomass_kg),
      # All below are weighted by biomass
      avg_depth       = weightedMean(avgdepth, w = biomass_kg, na.rm = T),
      avg_bot_temp    = weightedMean(bottemp, w = biomass_kg, na.rm = T),
      avg_sur_temp    = weightedMean(surftemp, w = biomass_kg, na.rm = T),
      avg_lat         = weightedMean(decdeg_beglat, w = biomass_kg, na.rm = T),
      avg_lon         = weightedMean(decdeg_beglon, w = biomass_kg, na.rm = T),
      depth_sd        = weightedSd(avgdepth, w = biomass_kg, na.rm = T),
      temp_sd         = weightedSd(bottemp, w = biomass_kg, na.rm = T),
      lat_sd          = weightedSd(decdeg_beglat, w = biomass_kg, na.rm = T),
      lon_sd          = weightedSd(decdeg_beglon, w = biomass_kg, na.rm = T),
      .groups = "drop") 
}

weighted_data <- grouped_center_bio(clean_survey, est_year, season) %>%
  mutate(decade = 10*est_year %/% 10)
```

```{r}
#| label: bsb
#| echo: false
#| warning: false
#| message: false

biomass <- weighted_data %>% 
  filter(comname == "black sea bass" & est_year >= 1970 & !season %in% c("Winter", "Summer")) 

landings <- garfo_landings %>% 
  filter(SPPNAME == "SEA BASS, BLACK" & YEAR >= 1970) %>% 
  select(PORT.NAME, STATE, YEAR, LANDED.LBS, VALUE) %>% 
  mutate(LANDED.LBS = parse_number(LANDED.LBS),
         VALUE = parse_number(VALUE))

fed_license <- garfo_clean %>% 
  pivot_longer(cols = BLACK_SEA_BASS_1:TILEFISH_D, names_to = "license", values_to = "count") %>% 
  filter(!count == 0 & license == "BLACK_SEA_BASS_1") 
 

me_license <- maine_licenses %>% 
  select(landings_number, license_year, cpc, cps) %>% 
  pivot_longer(cols = cpc:cps, names_to = "code", values_to = "count") %>%
  filter(!count == 0) %>%
  left_join(maine_codes) # in order to compare, would need state landings but I don't think those exist here...


```

### Average biomass and center of biomass based on NEFSC bottom trawl survey

```{r}
#| label: biomass
#| echo: false
#| message: false
#| warnging: false

ggplot(data = biomass) +
  geom_line(aes(x = est_year, y = avg_lat, color = season)) +
  theme_gmri()+
  scale_color_gmri() +
  ggtitle("Seasonal center of biomass") +
  xlab("Year") + ylab("Biomass-weighted average latitude")

biomass %>%
  group_by(est_year) %>% 
  summarise(avg_lat = mean(avg_lat)) %>% 
  ggplot() +
  geom_line(aes(x = est_year, y = avg_lat)) + 
  geom_smooth(aes(x = est_year, y = avg_lat), method = "lm", se = FALSE, color = "#07a3b7", linetype = 2, alpha = 0.7) +
  ylab("Biomass-weighted average latitude") +
  xlab("Year") +
  ggtitle("Annual center of biomass") +
  theme_gmri()

biomass %>% 
  group_by(est_year, season) %>%
  summarise(total_biomass = sum(total_biomass)) %>% 
  ggplot()+
  geom_col(aes(x = est_year, y = total_biomass, fill = season), position = "dodge") + #fill = "#07a3b7") +
  scale_fill_gmri(palette = "main") +
  theme_gmri() +
  ggtitle("Annual surveyed biomass") +
  ylab("Biomass (kg)") +
  xlab("Year")


biomass %>% 
  group_by(est_year) %>%
  summarise(annual_biomass = mean(total_biomass)) %>% 
  ggplot()+
  geom_line(aes(x = est_year, y = annual_biomass)) + #fill = "#07a3b7") +
  scale_fill_gmri(palette = "main") +
  theme_gmri() +
  ggtitle("Annual mean surveyed biomass") +
  ylab("Biomass (kg)") +
  xlab("Year")

```

### Federally reported landings

Dealer data provided by Greater Atlantic Regional Fisheries Office

```{r}
#| label: landings 
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-width: 11
#| fig-height: 9

landings$coast <- factor(landings$STATE, levels = c("ME", "NH", "MA", "RI", "CT", "NY", "NJ", "DE", "MD", "VA", "NC" ))

landings %>%
  group_by(STATE, YEAR, coast) %>% 
  summarise(total_volume = sum(LANDED.LBS)/1000000) %>%
  ggplot() +
  geom_col(aes(x = YEAR, y = total_volume), fill = "#07a3b7", alpha = 0.7) +
  facet_wrap(~coast, ncol = 3, scales = "free_y") +
  ylim(c(0,NA)) +
  ggtitle("Landed volume") +
  ylab("Total landings (million lbs)") +
  xlab("Year") +
  theme_gmri(strip.background = element_rect(fill = "transparent"),
              strip.text = element_text(color = "black")) 

landings %>%
  group_by(STATE, YEAR, coast) %>% 
  summarise(total_value = sum(VALUE)/1000000) %>%
  ggplot() +
  geom_line(aes(x = YEAR, y = total_value), color = "#057872") +
  facet_wrap(~coast, ncol = 3, scales = "free_y") +
  ylim(c(0, NA)) +
  ggtitle("Landed value") +
  ylab("Total landings (million $USD)") +
  xlab("Year") +
  theme_gmri(strip.background = element_rect(fill = "transparent"),
              strip.text = element_text(color = "black")) 

```

```{r}
#| label: totals
#| echo: false
#| message: false 
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 7
landings %>%
  group_by(YEAR) %>% 
  summarise(mean_volume = mean(LANDED.LBS)/1000000) %>%
  ggplot() +
  geom_col(aes(x = YEAR, y = mean_volume), fill = "#07a3b7", alpha = 0.7) + 
  geom_smooth(aes(x = YEAR, y = mean_volume), method = "lm", se = FALSE, linewidth = 1, color = "#535353", alpha = 0.8) +
  theme_gmri() +
  ggtitle("Average landed volume", subtitle = "Across all ports") +
  xlab("Year") + ylab("Landed pounds (million)") -> mean

landings %>%
  group_by(YEAR) %>% 
  summarise(total_volume = sum(LANDED.LBS)/1000000) %>%
  ggplot() +
  geom_col(aes(x = YEAR, y = total_volume), fill = "#07a3b7", alpha = 0.7) + 
  geom_smooth(aes(x = YEAR, y = total_volume), method = "lm", se = FALSE, linewidth = 1, color = "#535353", alpha = 0.8) +
  theme_gmri() + 
  ggtitle("Total landed volume", subtitle = "Across all ports") +
  xlab("Year") + ylab("Landed pounds (million)") -> total


patchwork::wrap_plots(total, mean, ncol = 2)

```

### Ex-vessel price of black sea bass

Despite a general upward trend in landed volume and value, the price per pound has dropped significantly in all regions.

```{r}
#| label: ex-vessel price
#| echo: false
#| message: false
#| warning: false

ex_vessel <- landings %>% 
  mutate(EX.VESSEL = LANDED.LBS/VALUE) 

ex_vessel %>% 
  group_by(STATE, YEAR) %>% 
  summarise(AVG.EX.VESSEL = mean(EX.VESSEL)) %>% 
  filter(!STATE == "ME") %>% # major outlier
  ggplot() + 
  geom_line(aes(x = YEAR, y = AVG.EX.VESSEL, color = STATE)) +
  ggtitle("Average ex-vessel price") + 
  xlab("Year") + ylab("Price per pound (USD)") + 
  theme_gmri() +
  scale_color_gmri()  

s_and_d <- ex_vessel %>% 
  group_by(YEAR) %>%
  summarise(AVG.LANDED.LBS = mean(LANDED.LBS/100000),
            AVG.EX.VESSEL  = mean(EX.VESSEL))
  
# ggplot(s_and_d) +
#   geom_line(aes(x = YEAR, y = AVG.LANDED.LBS), color = gmri_cols("ecosystems cyan")) + 
#   geom_line(aes(x = YEAR, y = AVG.EX.VESSEL),  color = gmri_cols("moss green")) +
#   scale_color_gmri() +
#   theme_gmri()

  
```

### Federal Licenses

There are two categories of federally issued black sea bass licenses.

-   **Category** 1: Commercial (moratorim)

    -   *Commercial (moratorium) permits have been managed under a limited entry system since 1996; no new moratorium permits are being issued.*

-   **Category 2**: Recreational charter/party

Given our focus on commercial harvest, we will focus only on **Category** 1.

```{r}
#| label: federal licenses
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-width: 11
#| fig-height: 9

# arrange north to south

fed_license$states <- factor(fed_license$PPST, levels = c("ME", "NH", "MA", "RI", "CT", "NY", "NJ", "DE", "MD", "VA", "NC" ))

fed_license %>% 
  group_by(PPST, AP_YEAR, states) %>% 
  filter(!PPST %in% c("FL", "GA", "PA", "WA")) %>% 
  summarise(total_count = sum(count)) %>%
  ggplot() +
  geom_col(aes(x = AP_YEAR, y = total_count), fill = gmri_cols("seafood purple")) +
  facet_wrap(~states, scales = "free_y", ncol = 3) +
  ggtitle("Commercial harvest licenses") +
  xlab("Year") + ylab("Total issued") +
  scale_fill_gmri() +
  theme_gmri(strip.background = element_rect(fill = "transparent"),
              strip.text = element_text(color = "black")) 

## Check that 646 licenses were issued in 2017 (re the noaa fisheries site)

# fed_license %>% 
#   filter(!PPST %in% c("FL", "GA", "PA", "WA")) %>% 
#   group_by(AP_YEAR) %>% 
#   summarise(total_count = sum(count)) # 669, figure this out

```
*If Commercial black sea pass permits have been under limited entry since 1996, why are there increases in the number of permits held in some years. Should they all be declining?*

## Concentration of licenses & landings 

```{r}
#| label: concentrations
#| echo: false
#| message: false
#| warning: false

library(rnaturalearth)

world <- ne_states(country = "united states of america", returnclass = "sf")

biomass %>%
  filter(est_year >= 2010) %>%
  group_by(est_year) %>% 
  summarise(avg_lat = mean(avg_lat),
            avg_lon = mean(avg_lon),
            avg_biomass = mean(avg_biomass)) -> recent
  
ggplot() +
  geom_sf(data = world)+
  coord_sf(xlim = c(-74, -70), ylim = c(38,42)) +
  geom_point(data = recent, aes(x = avg_lon, y = avg_lat, color = as.factor(est_year), size = avg_biomass)) +
  guides(size = guide_legend(title = "Average biomass (kg)"),
         color = guide_legend(title = "Year")) +
  ggtitle("Recent centers of biomass") +
  xlab("Lat") + ylab("Lon") +
  theme_gmri() +
  scale_color_gmri(palette = "mixed")
  
# Look at concentration (center of gravity) of landings in recent years
library(tidygeocoder)

landings %>% 
  geocode(city = PORT.NAME, state = STATE) -> landings

landings_conc <- landings %>% 
  group_by(YEAR) %>% 
  summarise(
    avg_lat = weightedMean(lat,  w = LANDED.LBS, na.rm = T),
    avg_lon = weightedMean(long, w = LANDED.LBS, na.rm = T)
  ) %>% 
  filter(YEAR >= 1996) %>% 
  mutate(decade = 10*YEAR %/% 10)

ggplot(landings_conc) +
  geom_line(aes(x = YEAR, y = avg_lat)) +
  ylim(c(36, 45)) +
  ggtitle("Concentration of landings", subtitle = "Since limited entry access") +
  xlab("Year") + ylab("Latitude") +
  theme_gmri()

# ggplot() +
#   geom_sf(data = world)+
#   coord_sf(xlim = c(-78, -66), ylim = c(36,45)) +
#   geom_point(data = landings_conc, aes(x = avg_lon, y = avg_lat)) +
#   theme_gmri()

# ggplot() +
#   geom_sf(data = world)+
#   coord_sf(xlim = c(-78, -66), ylim = c(36,45)) +
#   geom_point(data = (landings %>% filter(YEAR >= 1996 & LANDED.LBS > 10)), aes(x = long, y = lat, alpha = LANDED.LBS)) +
#   theme_gmri()

```
