---
title: "License Network Plots"
author: "cslovas"
format: 
  html:
    toc: true
    self-contained: true
editor: visual
---

Please review to the [License Networks](https://carlylovas.github.io/mesg-permits/R/license_networks.html) to understand which license code corresponds with which license type

```{r}
#| label:   load libraries & data 
#| echo:    false
#| message: false
#| warning: false

library(tidyverse)
library(tidygraph)
library(ggraph)
library(igraph)
library(here)
library(gmRi)
library(grid)

annual_graphs <- read_rds(here("Data", "network_graphs.rds"))

```

## Annual network plots

These plots include *all* data points, including years in which certain licenses may not have been issued

```{r}
#| label: annual graphs
#| echo: false
#| message: false
#| warning: false
#| column:  page
#| fig-width:  6
#| fig-height: 6
#| fig-align: center


gridExtra::marrangeGrob(annual_graphs$ggraph_circular, layout_matrix = matrix(1:1, ncol = 1, nrow = 1, byrow= TRUE), top = NULL)

```

## Edge density

Below is the time series of network edge density. The density of a graph is the ratio of the actual number of edges and the largest possible number of edges in the graph, assuming that no multi-edges are present. A network with higher density suggests more connections within the network. In the case of license networks, a more dense network would suggest more accesibility across fisheries and a less dense network would have less accessibility across fisheries. 

```{r}
#| label: edge density
#| echo: false
#| message: false
#| warning: false
#| column:  page
#| fig-width:  9
#| fig-height: 7
#| fig-align: center

annual_graphs %>%
  select(license_year, density) %>% 
  mutate(density = as.numeric(density)) %>%
  ggplot() +
  geom_line(aes(x = license_year, y = density)) +
  ggtitle("Edge density") +
  theme_gmri(plot.title = element_text(size = 10),
             axis.text  = element_text(size = 8),
             axis.title = element_blank())
              
```
