---
title: "License Networks"
author: "cslovas"
format: 
  html:
    toc: true
    self-contained: true
editor: visual
---

```{r}
#| label:   load libraries & data
#| echo:    false
#| message: false
#| warning: false

# Install & load the libraries
library(tidyverse)
library(tidygraph)
library(ggraph)
library(igraph)
library(here)
library(gmRi)
library(grid)
library(gt)

# Read in gomfish package from Joshua and Theresa
devtools::install_github("Social-Oceans-Lab/gomfish")
library(gomfish)

data(ind_lic_portfolio) 

ind_lic_portfolio <- ind_lic_portfolio %>% filter(!license_year == "2022") # unreliable

# Read in updated license codes 
updated_license_codes <- read.csv(here("Data", "updated_license_codes.csv")) %>%
  rename("license_type" = "code")

```

## Harvester licenses in Maine

To focus on fishery access and harvester adaptability, we've removed indigenous licenses, non-residential, and post-harvest (i.e., processors, wholesalers, etc) from the data set.

```{r}
#| label:   cleaning data
#| echo:    false
#| message: false
#| fig-height: 9

# Remove indigenous and non-harvester licenses
license_types <- names(ind_lic_portfolio %>%
                         select(!c(landings_number, license_year)))
license_types <- as_tibble(license_types) %>%
  rename("license_type" = "value") %>%
  left_join(updated_license_codes) %>%
  mutate(description = str_to_sentence(description))

license_types %>%
  filter(!group %in% c("Demo","Post-Harvest", "Non-resident", "Recreational")) %>%
  filter(!(str_starts(license_type, "ma"))) %>%
  filter(!(str_starts(license_type, "mi"))) %>%
  filter(!(str_starts(license_type, "nb"))) %>%
  filter(!(str_starts(license_type, "p")))  %>%
  filter(!license_type %in% c("st", "sts", "csw", "csws", "fhc", "lnc", "mws", "mw", "nfhc", "nfhs", "vh")) %>%
  drop_na() -> license_types # the one NA here is sut 

license_types %>% 
  gt() %>% 
  cols_hide(group) %>%
  cols_label(
    license_type   = md("**License Code**"),
    description = md("**License Type**"))

```

## Licenses Issued

Below is a time series depicting the total number of harvester licenses issued per year (line) and the number of licensed individuals per year (column)

```{r}
#| label:   total licenses issued
#| echo:    false
#| message: false
#| warning: false
#| column: page
#| fig-width:  8
#| fig-height: 6
#| fig-align: center

ind_lic_portfolio %>%
  select(license_year, license_types$license_type) %>% # landings_number) %>%
  pivot_longer(cols = 2:53, names_to = "license_type", values_to = "count") %>%
  # pivot_longer(cols = 3:147, names_to = "license_type", values_to = "count") %>%
  left_join(license_types) %>%
  group_by(license_year) %>%
  summarise(total = sum(count)) -> total_licenses

ind_lic_portfolio %>% 
  select(license_year, landings_number) %>%
  distinct() %>%
  group_by(license_year) %>%
  nest() %>%
  mutate(total_ind = map_dbl(data, function(x){nrow(x)})) %>%
  select(!data) %>%
  full_join(total_licenses) -> total_licenses


ggplot(data = total_licenses) +
  geom_col(aes(x = license_year, y = total_ind, fill = total_ind)) +
  geom_line(aes(x = license_year, y = total)) +
  ggtitle("Total Licenses Issued") +
  ylab("Total") + xlab("Year") +
  guides(fill = guide_colorbar(title = "Total Individuals", label = F)) +
  ylim(c(NA,20000)) +
  theme_gmri(plot.title = element_text(size = 10),
             axis.text  = element_text(size = 10),
             axis.title = element_text(size = 10))

```

Available licenses types change over time as new opportunities and management practices take shape and fisheries expand or close. Below is a time series depicting the amount of unique harvester license types available in a given year.

```{r}
#| label:   unique license types each year 
#| echo:    false
#| message: false
#| warning: false
#| column:  page
#| fig-width:  8
#| fig-height: 6
#| fig-align: center

ind_lic_portfolio %>% 
  select(license_year, license_types$license_type) %>%
  pivot_longer(cols = 2:53, names_to = "license_type", values_to = "count") %>%
  filter(!count == 0) %>% 
  select(license_year, license_type) %>% 
  distinct() %>%
  group_by(license_year) %>%
  nest() %>% 
  arrange(license_year) %>%
  mutate(unique_license_types = map_dbl(data, function(x){nrow(x)})) %>%
  rename("licenses" = "data")-> unique_licenses

ggplot(unique_licenses)+
  geom_line(aes(x = license_year, y = unique_license_types)) +
  ggtitle("Unique License Types") +
  theme_gmri(plot.title = element_text(size = 10),
             axis.text  = element_text(size = 10),
             axis.title = element_blank())
```

## Time series of license holdings

The total amount of licenses issued per year.

```{r}
#| label:   time series of license types
#| echo:    false
#| message: false
#| warning: false
#| column:  screen
#| fig-width:  9
#| fig-height: 9
#| fig-align: center

ind_lic_portfolio %>% 
  select(license_year, license_types$license_type) %>%
  pivot_longer(cols = 2:53, names_to = "license_type", values_to = "count") %>%
  left_join(license_types) %>% 
  #mutate(description = tolower(description)) %>%
  group_by(license_year, license_type, description) %>%
  summarise(total = sum(count)) %>%
  #filter(!total == 0) %>% 
  group_by(description) %>%
  nest() -> license_time_series

# plot
license_time_series %>% 
  #mutate(license = str_to_title(license)) %>%
  mutate(time_series = map2(data, description, function(x,y){
    out <- ggplot(data = x)+
      geom_line(aes(x = license_year, y = total)) +
      ylim(c(0, NA)) +
      xlim(c(1990, NA)) +
      theme_gmri(plot.title = element_text(size = 8),
                 axis.text  = element_text(size = 8),
                 axis.title = element_blank()) + 
      ggtitle(description)
    return(out)
  })) -> license_time_series

gridExtra::marrangeGrob(license_time_series$time_series, layout_matrix = matrix(1:12, ncol = 4, nrow = 3, byrow = TRUE), top = NULL,  left = textGrob(
  expression(bold("Total Annual")), rot = 90, gp = gpar(col = "black", fontsize = 10)))
```

## Average license holdings

Below is the average number of licenses held by an individual per year. Individuals who were not assigned a landings number (NA) were removed from the dataset before averages were calculated.

```{r}
#| label:   average license holdings
#| echo:    false
#| message: false
#| warning: false
#| column:  page
#| fig-width:  8
#| fig-height: 6
#| fig-align: center

# nrow(ind_lic_portfolio %>%
#   select(landings_number) %>%
#   distinct())

ind_lic_portfolio %>% 
  select(license_year, license_types$license_type, landings_number) %>%
  pivot_longer(cols = 2:53, names_to = "license_type", values_to = "count") %>%
  filter(!count == 0) %>%
  #filter(!is.na(landings_number)) %>%
  left_join(license_types) %>%
  group_by(license_year, landings_number) %>%
  distinct() %>%
  mutate(total = sum(count)) -> average_licenses
  
# average_licenses %>% filter(is.na(landings_number))

average_licenses %>%
  group_by(license_year) %>%
  nest() %>%
  mutate(average   = map_dbl(data, function(x){mean(x$total)})) %>%
         #total_ind = map_dbl(data, function(x){nrow(x)})) %>%
  select(!data) -> average_licenses

# (summary(average_licenses))

# summary(lm(total ~ license_year, data = average_licenses))

# gt(broom::glance(lm(average ~ license_year, data = average_licenses)))

average_licenses %>% 
  ggplot() +
  geom_point(aes(x = license_year, y = average)) +
  geom_line(aes(x = license_year, y = average), linewidth = 0.5) +
  stat_smooth(aes(x = license_year, y = average), method = "lm") +
  ylim(c(1,3)) +
  ggtitle("Average number of licenses held per individual") +
  theme_gmri(plot.title = element_text(size = 10),
             axis.text  = element_text(size = 10),
             axis.title = element_blank())
```

## Annual License Network Metrics

### Degree Centrality

"The *degree* of a node is the number of other nodes that single node is connected to.

Important nodes tend to have more connections to other nodes. Highly connected nodes are interpreted to have high degree centrality."

```{r}
#| label:      annual metrics
#| echo:       false
#| message:    false
#| warning:    false
#| column:     page
#| fig-width:  9
#| fig-height: 13
#| fig-align: center

annual_co_matrix <- ind_lic_portfolio %>%
  arrange(license_year) %>%
  #filter(!is.na(landings_number)) %>% # removing NA landings
  select(license_types$license_type, license_year) %>%
  group_by(license_year) %>%
  nest() %>% 
  left_join(unique_licenses) %>%
  select(license_year, data, licenses) %>% 
  mutate(co_matrix = map2(data, licenses, function(x, y){
    df <- x %>% select(y$license_type) # this filters the portfolio to feature only licenses issued in that year. If you want to run with all licenses, command out this line. 
    out <- crossprod(as.matrix(df)) 
    diag(out) <- 0
    return(out)
  })) %>%
  select(license_year, co_matrix) 

annual_metrics <- annual_co_matrix %>%
  mutate(graph_obj     = map(co_matrix, function(x){graph.adjacency(x, mode = "undirected", weighted = TRUE)}),
         degree        = map(graph_obj, function(x){degree(x)}),
         betweeness    = map(graph_obj, function(x){betweenness(x)}),
         density       = map(graph_obj, function(x){edge_density(x, loops = FALSE)}),
         evcent        = map(graph_obj, function(x){evcent(x)}))

metrics_series <- annual_metrics %>% 
  select(license_year, degree) %>% 
  unnest_longer(degree) %>% 
  group_by(degree_id) %>%
  rename("license_type" = "degree_id") %>%
  full_join(annual_metrics %>% 
              select(license_year, betweeness) %>% 
              unnest_longer(betweeness) %>% 
              group_by(betweeness_id) %>%
              rename("license_type" = "betweeness_id")) %>% 
  relocate("license_type", .after = "license_year") %>%
  nest() %>%
  left_join(license_types) %>%
  arrange(description)

degree_plots <- metrics_series %>% 
  mutate(time_series = map2(data, description, function(x, description){
    out <- ggplot(data = x)+
      geom_line(aes(x=license_year, y=degree)) +
      ylim(c(0, 50)) + 
      xlim(c(1990, 2021)) +
      theme_gmri(plot.title = element_text(size = 8),
                 axis.text  = element_text(size = 6),
                 axis.title = element_blank(),
                 strip.text = element_text(size = 8, color = "black"),
                 strip.background = element_blank()) +
      ggtitle(description)
    return(out)
  })) 

gridExtra::marrangeGrob(degree_plots$time_series, layout_matrix = matrix(1:15, ncol = 3, nrow = 5, byrow= TRUE), top = NULL, left = textGrob(
  expression(bold("Degree")), rot = 90, gp = gpar(col = "black", fontsize = 8)))
```

### Betweeness

"\[*Betweenness*\] ranks the nodes based on the flow of connections through the network. Importance is demonstrated through high frequency of connection with multiple other nodes. Nodes with high levels of betweenness tend to serve as a bridge for multiple sets of other important nodes." [*Source*](https://bookdown.org/jdholster1/idsr/network-analysis.html#advanced-case-study)

License times with high and/or increasing betweeness values may represent important "fringe" fisheries; fisheries that have a low cost of entry, use existing equipment and/or are harvested by hand. These "fringe" fisheries may serve as additional or supplemental means of income.

```{r}
#| label:      betweeness
#| echo:       false
#| warning:    false
#| column:     page
#| fig-width:  9
#| fig-height: 13
#| fig-align: center

betweeness_plots <- metrics_series %>% 
  mutate(time_series = map2(data, description, function(x, description){
    out <- ggplot(data = x)+
      geom_line(aes(x=license_year, y=betweeness)) +
      ylim(c(0, 150)) + 
      xlim(c(1990, 2021)) +
      theme_gmri(plot.title = element_text(size = 8),
                 axis.text  = element_text(size = 6),
                 axis.title = element_blank(),
                 strip.text = element_text(size = 8, color = "black"),
                 strip.background = element_blank()) +
      ggtitle(description)
    return(out)
  })) 

gridExtra::marrangeGrob(betweeness_plots$time_series, layout_matrix = matrix(1:15, ncol = 3, nrow = 5, byrow= TRUE), top = NULL, left = textGrob(
  expression(bold("Degree")), rot = 90, gp = gpar(col = "black", fontsize = 8)))
```

## Annual Network Plots

Density time series and network plots are available [here](https://carlylovas.github.io/mesg-permits/R/license_network_plots.html)

```{r}
#| label: network plots
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 8
#| fig-align: center

annual_graphs <- annual_metrics %>%
  mutate(ggraph = map2(graph_obj, license_year, function(x, y){
    x <- as_tbl_graph(x)
    x %>% 
      activate(nodes) %>%
      mutate(community = as.character(group_louvain())) -> x
    out <- ggraph(x, layout = "fr") + 
      geom_node_point(size = 7, alpha = 0.5, show.legend = FALSE) +
      geom_edge_link(aes(width = weight), alpha = 0.5, show.legend = FALSE) + 
      scale_edge_width(range = c(0.2, 2)) +
      ggtitle(license_year) +
      geom_node_text(aes(label = name), repel = TRUE, color = "black", max.overlaps = 100) +
      theme_graph()   
  })) %>%
  mutate(ggraph_circular = map2(graph_obj, license_year, function(x, y){
    x <- as_tbl_graph(x)
    x %>% 
      activate(nodes) %>%
      mutate(community = as.character(group_louvain())) -> x
    out <- ggraph(x, layout = "linear", circular = TRUE) + 
      geom_node_point(size = 7, alpha = 0.5, show.legend = FALSE) +
      geom_edge_link(aes(width = weight), alpha = 0.5, show.legend = FALSE) + 
      scale_edge_width(range = c(0.2, 2)) +
      ggtitle(license_year) +
      geom_node_text(aes(label = name), repel = TRUE, color = "black", max.overlaps = 100) +
      theme_graph()   
  })) %>%
  select(license_year, density, ggraph, ggraph_circular)


write_rds(annual_graphs, here("Data", "network_graphs.rds"))

# make different markdown for network plots; too big for github
```
