---
title: "Species Networks"
author: "cslovas"
format: 
  html:
    toc: true
    self-contained: true
editor: visual
---

```{r, message = FALSE}
#| label: load packages & data
#| echo: false
#| warning: false
library(tidyverse)
library(janitor)
library(tidygraph)
library(ggraph)
library(igraph)
library(patchwork)
library(gmRi)
library(grid)
library(here)
library(gt)

# Read in gomfish package from Joshua and Theresa
devtools::install_github("Social-Oceans-Lab/gomfish")
library(gomfish)

data(ind_lic_portfolio) 

ind_lic_portfolio <- ind_lic_portfolio %>% filter(!license_year == "2022") # unreliable

# Read in updated license codes 
updated_license_codes <- read.csv(here("Data", "updated_license_codes.csv")) %>%
  rename("license_type" = "code")

```

## Species Grouping

To correlate species distribution changes with changes in license holdings in Maine, license types have been grouped by target species as described by Maine State Legislation.

```{r}
#| label: license types
#| echo: false
#| message: false
#| warning: false
#| column: page

# Remove indigenous and non-harvester licenses
license_types <- names(ind_lic_portfolio %>%
                         select(!c(landings_number, license_year)))
license_types <- as_tibble(license_types) %>%
  rename("license_type" = "value") %>%
  left_join(updated_license_codes) %>%
  mutate(description = str_to_sentence(description))

license_types %>%
  filter(!group %in% c("Demo","Post-Harvest", "Non-resident", "Recreational")) %>%
  filter(!(str_starts(license_type, "ma"))) %>%
  filter(!(str_starts(license_type, "mi"))) %>%
  filter(!(str_starts(license_type, "nb"))) %>%
  filter(!(str_starts(license_type, "p")))  %>%
  filter(!license_type %in% c("st", "sts", "csw", "csws", "fhc", "lnc", "mws", "mw", "nfhc", "nfhs", "vh")) %>%
  drop_na() -> license_types # the one NA here is sut 

license_types %>% 
  gt() %>% 
  cols_hide(license_type) %>%
  cols_label(
    group   = md("**Target Species**"),
    description = md("**License Type**"))
```

## License Portfolio by Species

Using Theresa Burnham's code for building the individual license portfolio, license types were grouped by target species. A binary matrix was created, whereas 1 denotes an individual license holding for a corresponding year and 0 denotes no license held. With this co-occurrence matrix, we can quantify which license types were held together over the past 30 years and how those holdings have changed.

```{r, message=FALSE}
#| label: license portfolio
#| echo: false
#| message: false

ind_lic_portfolio %>%
  select(license_year,landings_number, license_types$license_type) %>%
  pivot_longer(cols = 3:52, names_to = "license_type", values_to = "licensed") %>%
  filter(licensed == 1) %>%
  left_join(license_types) %>%
  mutate(row = row_number()) %>%
  select(license_year, landings_number, licensed, group, row) %>%
  pivot_wider(names_from=group,values_from=licensed,names_expand=TRUE,values_fill = list(licensed=0)) %>%
  select(-row)%>%
  group_by(landings_number,license_year)%>%
  summarise(across(everything(),sum))-> species_portfolio

```

### Targeted Species
```{r}
#| label:   targeted species
#| echo:    false
#| message: false
#| warning: false
#| column:  page
#| fig-width:  8
#| fig-height: 6
#| fig-align: center

species_portfolio %>% 
  pivot_longer(cols = 3:21, names_to = "species", values_to = "count") %>%
  filter(count == 1) %>%
  ungroup(landings_number) %>%
  select(license_year, species) %>%
  distinct() %>%
  group_by(license_year) %>%
  nest() %>%
  arrange(license_year) %>%
  mutate(num_targeted_spp = map_dbl(data, function(x){nrow(x)})) %>%
  rename("species" = "data") -> targeted_species

ggplot(targeted_species)+
  geom_line(aes(x = license_year, y = num_targeted_spp)) +
  ggtitle("Number of Target Species") +
  theme_gmri(plot.title = element_text(size = 10),
             axis.text  = element_text(size = 10),
             axis.title = element_blank())
```

```{r}
#| label: species proportion
#| echo: false
#| message: false
#| warning: false
#| column: page
#| fig-align: center
#| fig-height: 6
#| fig-width: 8

species_portfolio %>% 
  pivot_longer(cols = 3:21, names_to = "species", values_to = "count") %>%
  group_by(license_year, species) %>%
  summarise(count = sum(count)) %>%
  group_by(license_year) %>%
  nest() %>%
  mutate(total = map_dbl(data, function(x){sum(x$count)})) %>%
  unnest(data) %>%
  group_by(license_year, species) %>%
  summarise(proportion = count/total) -> species_proportions


ggplot(data = species_proportions, aes(x = license_year, y = proportion, fill = species)) +
  geom_col(position = "stack") +
  theme_gmri(plot.title = element_text(size = 10),
             axis.text  = element_text(size = 10),
             axis.title = element_text(size = 8)) +
  xlab("Year") + ylab("Proportion") + ggtitle("Proportion of Targeted Species") +
  scale_fill_viridis_d() +
  guides(fill = guide_legend(nrow=2))
```

```{r}
#| label: average number of species targeted
#| echo: false
#| message: false
#| warning: false
#| column: page
#| fig-align: center
#| fig-height: 6
#| fig-width: 8


species_portfolio %>% 
  pivot_longer(cols = 3:21, names_to = "species", values_to = "count") %>%
  filter(count == 1) %>%
  group_by(license_year, landings_number) %>% 
  nest() %>%
  mutate(num_spp = map_dbl(data, function(x){nrow(x)})) %>% 
  select(landings_number, license_year, num_spp) %>%
  group_by(license_year) %>%
  summarise(average_num_spp = mean(num_spp)) -> avg_targeted_spp

ggplot(avg_targeted_spp) +
  geom_point(aes(x = license_year, y = average_num_spp)) +
  geom_line(aes(x = license_year, y = average_num_spp), linewidth = 0.5) +
  stat_smooth(aes(x = license_year, y = average_num_spp), method = "lm") +
  ylim(c(1,2)) +
  ggtitle("Average number of targeted/harvested species per individual") +
  theme_gmri(plot.title = element_text(size = 10),
             axis.text  = element_text(size = 10),
             axis.title = element_blank())


```
## Time Series of Targeted Species
The plots below show the total number of licenses issued per year for a particular target species.

```{r}
#| label:   time series of target species
#| echo:    false
#| message: false
#| warning: false
#| column:  screen
#| fig-width:  9
#| fig-height: 9
#| fig-align: center

ind_lic_portfolio %>% 
  select(license_year, license_types$license_type) %>%
  pivot_longer(cols = 2:51, names_to = "license_type", values_to = "count") %>%
  left_join(license_types) %>%
  select(license_year, group, count) %>%
  group_by(license_year, group) %>%
  summarise(count = sum(count)) %>%
  group_by(group) %>% 
  nest() %>%
  mutate(time_series = map2(data, group, function(x,y){
    out <- ggplot(data = x)+
      geom_line(aes(x = license_year, y = count)) +
      ylim(c(0, NA)) +
      xlim(c(1990, NA)) +
      theme_gmri(plot.title = element_text(size = 8),
                 axis.text  = element_text(size = 8),
                 axis.title = element_blank()) + 
      ggtitle(group)
    return(out)
  })) -> time_series

gridExtra::marrangeGrob(time_series$time_series, layout_matrix = matrix(1:12, ncol = 4, nrow = 3, byrow = TRUE), top = NULL,  left = textGrob(
  expression(bold("Total Annual")), rot = 90, gp = gpar(col = "black", fontsize = 10)))



```
## Annual Species Metrics
### Degree centrality

"The *degree* of a node is the number of other nodes that single node is connected to.

Important nodes tend to have more connections to other nodes. Highly connected nodes are interpreted to have high degree centrality."

```{r}
#| label:      annual metrics
#| echo:       false
#| message:    false
#| warning:    false
#| column:     page
#| fig-width:  9
#| fig-height: 13
#| fig-align: center

annual_co_matrix <- species_portfolio %>%
  arrange(license_year) %>%
  select(!landings_number) %>%
  group_by(license_year) %>%
  nest() %>% 
  left_join(targeted_species) %>%
  select(license_year, data, species) %>% 
  mutate(co_matrix = map2(data, species, function(x, y){
    df <- x %>% select(y$species) # this filters the portfolio to feature only species harvested in that year. If you want to run with all species, command out this line
    out <- crossprod(as.matrix(df)) 
    diag(out) <- 0
    return(out)
  })) %>%
  select(license_year, co_matrix) 

annual_metrics <- annual_co_matrix %>%
  mutate(graph_obj     = map(co_matrix, function(x){graph.adjacency(x, mode = "undirected", weighted = TRUE)}),
         degree        = map(graph_obj, function(x){degree(x)}),
         betweeness    = map(graph_obj, function(x){betweenness(x)}),
         density       = map(graph_obj, function(x){edge_density(x, loops = FALSE)}),
         evcent        = map(graph_obj, function(x){evcent(x)}))

metrics_series <- annual_metrics %>% 
  select(license_year, degree) %>% 
  unnest_longer(degree) %>% 
  group_by(degree_id) %>%
  rename("species" = "degree_id") %>%
  full_join(annual_metrics %>% 
              select(license_year, betweeness) %>% 
              unnest_longer(betweeness) %>% 
              group_by(betweeness_id) %>%
              rename("species" = "betweeness_id")) %>% 
  relocate("species", .after = "license_year") %>%
  nest() %>%
  arrange(species)

degree_plots <- metrics_series %>% 
  mutate(time_series = map2(data, species, function(x, description){
    out <- ggplot(data = x)+
      geom_line(aes(x=license_year, y=degree)) +
      ylim(c(0, 40)) + 
      xlim(c(1990, 2021)) +
      theme_gmri(plot.title = element_text(size = 8),
                 axis.text  = element_text(size = 6),
                 axis.title = element_blank(),
                 strip.text = element_text(size = 8, color = "black"),
                 strip.background = element_blank()) +
      ggtitle(species)
    return(out)
  })) 

gridExtra::marrangeGrob(degree_plots$time_series, layout_matrix = matrix(1:12, ncol = 3, nrow = 4, byrow= TRUE), top = NULL, left = textGrob(expression(bold("Degree")), rot = 90, gp = gpar(col = "black", fontsize = 8)))
```
### Betweeness
"\[*Betweenness*\] ranks the nodes based on the flow of connections through the network. Importance is demonstrated through high frequency of connection with multiple other nodes. Nodes with high levels of betweenness tend to serve as a bridge for multiple sets of other important nodes." [*Source*](https://bookdown.org/jdholster1/idsr/network-analysis.html#advanced-case-study)

```{r}
#| label:      annual metrics pt 2
#| echo:       false
#| message:    false
#| warning:    false
#| column:     page
#| fig-width:  9
#| fig-height: 13
#| fig-align: center

betweeness_plots <- metrics_series %>% 
  mutate(time_series = map2(data, species, function(x, description){
    out <- ggplot(data = x)+
      geom_line(aes(x=license_year, y=betweeness)) +
      ylim(c(0, 50)) + 
      xlim(c(1990, 2021)) +
      theme_gmri(plot.title = element_text(size = 8),
                 axis.text  = element_text(size = 6),
                 axis.title = element_blank(),
                 strip.text = element_text(size = 8, color = "black"),
                 strip.background = element_blank()) +
      ggtitle(species)
    return(out)
  })) 

gridExtra::marrangeGrob(betweeness_plots$time_series, layout_matrix = matrix(1:12, ncol = 3, nrow = 4, byrow= TRUE), top = NULL, left = textGrob(expression(bold("Degree")), rot = 90, gp = gpar(col = "black", fontsize = 8)))
```


## Annual Network Plots

Similar to plotting the license holdings as a network over time, we plot the target species pertaining to these permits over time. This makes for a more easily interpretable graph, as well as increased potential for correlating license changes to species distribution changes

```{r}
#| label: species networks
#| echo: false
#| message:    false
#| warning:    false
#| column:     page
#| fig-width:  6
#| fig-height: 6
#| fig-align: center


# Network plots
annual_species_graphs <- annual_metrics %>%
  mutate(ggraph = map2(graph_obj, license_year, function(x, y){
    x <- as_tbl_graph(x)
    x %>% 
      activate(nodes) %>%
      mutate(community = as.character(group_louvain())) -> x
    out <- ggraph(x, layout = "fr") + 
      geom_node_point(size = 7, alpha = 0.5, show.legend = FALSE) +
      geom_edge_link(aes(width = weight), alpha = 0.5, show.legend = FALSE) + 
      scale_edge_width(range = c(0.2, 2)) +
      ggtitle(license_year) +
      geom_node_text(aes(label = name), repel = TRUE, color = "black", max.overlaps = 100) +
      theme_graph()   
  })) %>%
  mutate(ggraph_circular = map2(graph_obj, license_year, function(x, y){
    x <- as_tbl_graph(x)
    x %>% 
      activate(nodes) %>%
      mutate(community = as.character(group_louvain())) -> x
    out <- ggraph(x, layout = "linear", circular = TRUE) + 
      geom_node_point(size = 7, alpha = 0.5, show.legend = FALSE) +
      geom_edge_link(aes(width = weight), alpha = 0.5, show.legend = FALSE) + 
      scale_edge_width(range = c(0.2, 2)) +
      ggtitle(license_year) +
      geom_node_text(aes(label = name), repel = TRUE, color = "black", max.overlaps = 100) +
      theme_graph()   
  })) %>%
  select(license_year, ggraph, ggraph_circular)

gridExtra::marrangeGrob(annual_species_graphs$ggraph_circular, layout_matrix = matrix(1:1, ncol = 1, nrow = 1, byrow= TRUE), top = NULL)
```

## Limited Entry Fisheries

According to [Maine State Legislation](http://www.mainelegislature.org/legis/statutes/12/title12ch0sec0.html), there are 6 limited entry species including lobster, which is limited by harvest zone. Most species became limited entry in response to a fishery collapse, while some aim to avoid collapse. The most recent limited entry fishery is menhaden, which was passed by the legislature in 2023.

```{r}
#| message: false
#| label: limited entry
#| echo: false

# fishery_entry <- license_types %>% 
#   select(group) %>% 
#   distinct() %>%
#   filter(group %in% c("Scallop", "Sea urchin", "Sea cucumber", "Menhaden", "Elver", 
#                       "Lobster")) %>%
#   mutate(entry      = "limited",
#          entry_year = c("2011", "2015", "2023", "2009", "2005", "2007"))
# 
# license_types %>% 
#   select(group) %>%
#   distinct() %>%
#   filter(!group %in% c("Scallop", "Sea urchin", "Sea cucumber", "Menhaden", "Elver", 
#                         "Lobster")) %>%
#   mutate(entry = "open",
#          entry_year = NA) %>% 
#   full_join(fishery_entry) -> fishery_entry
# 
# fishery_entry %>% 
#   filter(entry == "limited") %>% 
#   left_join(license_proportions) %>% 
#   mutate(entry_year = as.numeric(entry_year)) %>%
#   ggplot() +
#   geom_line(aes(x  = license_year, y = license_count))  +
#   geom_vline(aes(xintercept = entry_year), color = "#EA4D12", linetype = 2) + 
#   facet_wrap(~group, scales = "free_y") +
#   xlab("Year") + ylab("Number of Licenses Issued") +
#   theme_gmri() 
```
